---
title: "Week07_Homework"
author: "Kassandra Jensen"
date: "2024-10-13"
output: 
   html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE, 
                      message = FALSE)

```

## Directions
You need to make a map using RMarkdown with one of the following datasets from past Tidy Tuesdays. Make sure, as usual, that you scripts and output are saved appropriately.

I will be using the Tidy Tuesday data set: Cats on the move!

# Load Libraries
```{r}
library(tidytuesdayR)
library(ggmap)
library(tidyverse)
library(here)
library(ggspatial)
```

# Load data: Cats on the move!
The data and code below were provided [here.](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-01-31/readme.md)

```{r}
tuesdata <- tidytuesdayR::tt_load('2023-01-31')
tuesdata <- tidytuesdayR::tt_load(2023, week = 5)

cats_uk <- tuesdata$cats_uk
cats_uk_reference <- tuesdata$cats_uk_reference

glimpse(cats_uk)
glimpse(cats_uk_reference)
```

# Join data
I wanted to join the data to have all the information in one place, especially the actual cat name as well as the tag ID. 
```{r}
## Join both data sets by the column tag_id.
cats_uk_full <- full_join(cats_uk, cats_uk_reference, by = "tag_id")

## View to make sure it was joined properly.
view(cats_uk_full)  
```

# Filter data
I filtered the data so only a two cats would be shown. This way we can track their movements more closely, rather than looking across the entire United Kingdom.
```{r}
## Select cats Maverick and Bear.
selected_cats <- c("Maverick", "Bear")

# Filter the dataset to only include the selected cats.
cats_selected <- cats_uk_full %>% 
  filter(animal_id %in% selected_cats)

## View to make sure only the specified cats were chosen.
view(cats_selected)
```

# Make Map
## Get map location
```{r}
## Need to manually set location because UK map was too general. Pulled point from one of the cats.
UK_map <- get_map(location = c(lon = -5.081133, lat = 50.14673), zoom = 15, maptype = "hybrid")
ggmap(UK_map)
```

## Plot Maverick and Bear on the map. 
```{r}
ggmap(UK_map) +
  geom_point(data = cats_selected,                     ## select data set for points
             aes(x = location_long,                    ## longitude
                 y = location_lat,                     ## latitude
                 color = as.factor(animal_id)),        ## change color by each cat
             size = 2) +                               ## change to best looking size
  geom_path(data = cats_selected %>% arrange(animal_id, timestamp),    ## group by cat, connect in order of cat movement 
            aes(x = location_long,                     ## longitude
                y = location_lat,                      ## latitude
                group = animal_id,                     ## group by each cat
                color = as.factor(animal_id)),         ## color by each cat
            size = 0.5, lineend = "round") +           ## fix line aesthetics
  labs(x = "Longitude",                                ## fix labels and add a cheeky title
       y= "Latitude",
       color = "Cat Names",
       title="Whiskers in the Wild: Tracking Two Cats at Gyllyngvase Beach",
       subtitle = "A Journey Through the Eyes of Coastal Cats") +
  theme_minimal()

ggsave(here("Week07","Output","Homework", "Homework07_CatMap.png")) ## Save plot
```

