---
title: "Week16Homework"
author: "Kassandra Jensen"
date: "2024-12-05"
output: 
  html_document:
    toc: true
    toc_float: true
---

# Directions
Bring in all 4 files and calculate the mean and standard deviation of both temperature (Temp.C) and light (Intensity.lux) for each tide pool. Use both a for loop and map() functions in your script. (Basically, do it twice).

I am going to break it down step by step like we did in class because it's easier for me to understand that way.

# Libraries
```{r}
library(tidyverse)
library(here)
```

# Loop 
```{r}
#test out to make sure it will read at least one file:
datadata<-read_csv(here("Week15", "Homework","data", "TP1.csv"))
glimpse(datadata)
colnames(datadata)

## Establish the pathway
data_path<-here("Week15", "Homework","data")
            
## Find everything in this folder ending in ".csv"
homework_files <- dir(path = data_path, pattern = ".csv")
homework_files
```
## Make empty data frame
```{r}
homework_data<-tibble(filename =  rep(NA, length(homework_files)),  # column name for pool
                   mean_temp = rep(NA, length(homework_files)), # column name for the mean temperature
                   mean_lux = rep(NA, length(homework_files)), # column name for the mean light
                   sd_temp = rep(NA, length(homework_files)), # column names standard deviation of temp
                   sd_lux = rep(NA, length(homework_files)) # column name for standard deviation of lux
                   ) 
homework_data
```
## Read file to see if this works
```{r}
raw_HW_data<-read_csv(paste0(data_path,"/",homework_files[1])) # test by reading in the first file and see if it works
head(raw_HW_data)
```

## Calculate mean
```{r}
mean_temp_HW <-mean(raw_HW_data$Temp.C, na.rm = TRUE) # calculate a mean of the one we tested out
mean_temp_HW
```
## Make simple loop
```{r}
for (i in 1:length(homework_files)){ # loop over 1:4 the number of files
}
```

## Add in the loop over the raw data
```{r}
for (i in 1:length(homework_files)){ # loop over 1:4 the number of files 
raw_HW_data<-read_csv(paste0(data_path,"/",homework_files[i])) ## read the file like we made earlier
#glimpse(raw_HW_data)  ## glimpse to make sure it looks okay
}
```
## Add in filename for each row
```{r}
for (i in 1:length(homework_files)){ # loop over 1:4 the number of files 
raw_HW_data<-read_csv(paste0(data_path,"/",homework_files[i])) ## create path to read all the files
homework_data$filename[i]<-homework_files[i]                  ## 
} 
homework_data
```

## Finished loop!
```{r}
for (i in 1:length(homework_files)){ # loop over 1:4 the number of files 
  raw_HW_data<-read_csv(paste0(data_path,"/",homework_files[i]))  ## read in data and include in loop
  homework_data$filename[i]<-homework_files[i]                          ## add file names for each row
  homework_data$mean_temp[i]<-mean(raw_HW_data$Temp.C, na.rm =TRUE)    ## calculate mean temp and assign name
  homework_data$mean_lux[i]<-mean(raw_HW_data$Intensity.lux, na.rm =TRUE)  ## calculate mean temp and assign name
  homework_data$sd_temp[i]<-sd(raw_HW_data$Temp.C, na.rm =TRUE)  ## calculate sd temp and assign name
  homework_data$sd_lux[i]<-sd(raw_HW_data$Intensity.lux, na.rm =TRUE) ## calculate sd lux and assign name
} 
homework_data ## check to make sure it worked!
```

# Map function
Okay, now we do the same thing but using the map function instead of a loop.
## Get files
I'm using the same path as before.
```{r}
data_path<-here("Week15", "Homework","data")
homework_files <- dir(path = data_path, pattern = ".csv", full.names = TRUE) ## Had to make full.names= True, maybe bc my working directory isn't the same as my data directory?

homework_files
```
## Read in the files using map
```{r}
HW_data<- homework_files %>%
  set_names()%>% # set's the id of each list to the file name
  map_df(read_csv,.id = "filename") # map everything to a dataframe and put the id in a column called filename
HW_data
```
## Calculate means and sds
```{r}
HW_data<-homework_files %>%
  set_names()%>% # set's the id of each list to the file name
  map_df(read_csv,.id = "filename") %>% # map everything to a dataframe and put the id in a column called filename
  group_by(filename) %>%
  summarise(mean_temp = mean(Temp.C, na.rm = TRUE),
            mean_lux = mean(Intensity.lux,na.rm = TRUE),
            sd_temp = sd(Temp.C, na.rm = TRUE),
            sd_lux = sd(Intensity.lux,na.rm = TRUE)
            )%>%
  mutate(filename = basename(filename))  ## Because I used row.names = true, it gave me the whole path to these files and was really ugly so i used this code to extract only the basename of the file!
HW_data
```

















