---
title: "Tidy Tuesday 06"
author: "Kassandra Jensen"
date: "2024-12-02"
output: 
  prettydoc::html_pretty:
    theme: hpstr
    toc: yes
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = TRUE) 
```
***
# Libraries
```{r}
library(tidyverse)
library(gganimate)
library(prettydoc)
```
***
# Data
```{r}
A64_traffic <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-12-03/A64_traffic.csv')
```
***
# Data manipulation
Made the data frame longer, so I could graph the counts. 
```{r}
A64_traffic_long <- A64_traffic %>%
  pivot_longer(cols = '0 - 520 cm':'1160+ cm',
               names_to = "car_length",
               values_to = "count_length") %>%
  pivot_longer(cols = '0 - 10 mph':'80+ mph',
               names_to = "speed",
               values_to = "count_speed")
```
***
# Plot
This plot helped to get an initial understanding of the data and any trends.
```{r}
ggplot(A64_traffic_long, aes(x = speed, y = count_speed, fill = car_length)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Distribution of Vehicle Speeds",
    x = "Speed Category (mph)",
    y = "Count of Vehicles",
    fill = "Car Length (cm)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
***
# Final Plot
I wanted to see if there were any big changes overtime. It seems like Fridays always had the most cars on the road. 


I had to comment out the animation because it wouldn't knit. Sorry!
```{r}
animated_plot <- ggplot(A64_traffic_long, aes(x = speed, y = count_speed, fill = car_length)) + ## make plot
  geom_bar(stat = "identity", position = "stack") +                                             ## bar plot + stack keeps the different groups stacked on one another
  labs(
    title = "Distribution of Vehicle Speeds Over Time",                                         ## Labels
    subtitle = "Date: {May 1 - May 31, 2024}",
    x = "Speed Category (mph)",
    y = "Count of Vehicles",
    fill = "Car Length (cm)"
  ) + 
  theme_minimal() +                                                                             ## minimal theme
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  transition_time(as.Date(`Report Date`)) +                                                     ## animate over the report date
  ease_aes("linear")                                                                            ## makes it have smooth transitions


# Make the plot render a little bit slower to see more easiry
#animate(animated_plot, 
#        width = 800, 
#        height = 600, 
#        fps = 5,                                                                                ## fewer frames per second 
#        duration = 20)                                                                          ## total animation duration
```
***
# Save animation
I had to comment this out as well to get it to knit. Sorry about that. 
```{r}
# animate(animated_plot, width = 800, height = 600, fps = 5)
# anim_save("animated_traffic_plot.gif")
```

# New things I learned:
I learned how to change the R markdown theme using the *prettydoc* package. There are a ton of other themes to use. I found this one on this [website](https://rpubs.com/ranydc/rmarkdown_themes#:~:text=To%20apply%20a%20theme%20to,the%20beginning%20of%20your%20document).  

