---
title: "Tidy Tuesday Week 12"
author: "Kassandra Jensen"
date: "2024-11-14"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Load libraries
```{r}
library(here)
library(tidyverse)
```

# Read in data
```{r}
countries <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-12/countries.csv')
country_subdivisions <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-12/country_subdivisions.csv')
former_countries <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-12/former_countries.csv')

head(countries)
head(country_subdivisions)
head(former_countries)
```
# Data Manipulations
Extract year from dates and get counts for each year.
```{r}
withdrawals_per_year <- former_countries %>%
  mutate(year = substr(date_withdrawn, 1, 4)) %>%                  ## Extract the first 4 characters (in this case- year)
  group_by(year) %>%                                               ## Group by year
  summarise(count = n()) %>%                                       ## get nuber of occurences
  arrange(year)

withdrawals_per_year$year <- as.numeric(withdrawals_per_year$year) ## make year numeric
 
```

# Graph
Making a graph showing number of countries that have been withdrawn over time.
This is super basic but I had *no inspiration* for this data set and I did NOT want to make a map.
```{r}
ggplot(withdrawals_per_year, aes(x = year, y = count)) +           ## data, x and y
   geom_bar(stat = "identity", fill = "darkblue") +               ## use geom bar
  labs(
    title = "Number of Countries Withdrawing Per Year",           ## title and labels
    x = "Year",
    y = "Number of Countries"
  ) +
  theme_minimal() +                                               ## minimal theme
  coord_flip()                                                    ## flip axis

ggsave("Withdrawals.png")
```

# New thing I learned:
I learned how to fold code chunks in R markdown! I makes it so that the code can be hidden and you only get outputs. So simple! I really appreciate it because sometimes all you care about it the output. 



