---
title: "Week 11 Tidy Tuesday"
author: "Kassandra Jensen"
date: "2024-11-07"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Week 11 Tidy Tuesday: Democracy Data

# Load libraries
```{r}
library(democracyData)
library(tidyverse)
library(patchwork)
```

# Read in Data
This data comes from the tidytuesday data for November 11, 2024 and can be found [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2024/2024-11-05/readme.md).
There is data manipulation/ clean-up that was suggested on the repo for this data set which is what you see below.
```{r}
democracy_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-05/democracy_data.csv')

## Clean up data
## Taken rom repository
democracy_data <-
  democracyData::pacl_update |> 
  janitor::clean_names() |> 
  dplyr::select(
    "country_name" = "pacl_update_country",
    "country_code" = "pacl_update_country_isocode",
    "year",
    "regime_category_index" = "dd_regime",
    "regime_category" = "dd_category",
    "is_monarchy" = "monarchy",
    "is_commonwealth" = "commonwealth",
    "monarch_name",
    "monarch_accession_year" = "monarch_accession",
    "monarch_birthyear",
    "is_female_monarch" = "female_monarch",
    "is_democracy" = "democracy",
    "is_presidential" = "presidential",
    "president_name",
    "president_accesion_year" = "president_accesion",
    "president_birthyear",
    "is_interim_phase" = "interim_phase",
    "is_female_president" = "female_president",
    "is_colony" = "colony",
    "colony_of",
    "colony_administrated_by",
    "is_communist" = "communist",
    "has_regime_change_lag" = "regime_change_lag",
    "spatial_democracy",
    "parliament_chambers" = "no_of_chambers_in_parliament",
    "has_proportional_voting" = "proportional_voting",
    "election_system",
    "lower_house_members" = "no_of_members_in_lower_house",
    "upper_house_members" = "no_of_members_in_upper_house",
    "third_house_members" = "no_of_members_in_third_house",
    "has_new_constitution" = "new_constitution",
    "has_full_suffrage" = "fullsuffrage",
    "suffrage_restriction",
    "electoral_category_index" = "electoral",
    "spatial_electoral",
    "has_alternation" = "alternation",
    "is_multiparty" = "multiparty",
    "has_free_and_fair_election" = "free_and_fair_election",
    "parliamentary_election_year",
    "election_month" = "election_month_year",
    "has_postponed_election" = "postponed_election"
  ) |>
  dplyr::mutate(
    election_month = dplyr::na_if(.data$election_month, "?")
  ) |> 
  tidyr::separate_wider_regex(
    "election_month",
    patterns = c(
      election_month = "\\D+",
      election_year = "\\d{4}$"
    ),
    too_few = "align_start"
  ) |> 
  dplyr::mutate(
    electoral_category = dplyr::case_match(
      .data$electoral_category_index,
      0 ~ "no elections",
      1 ~ "single-party elections",
      2 ~ "non-democratic multi-party elections",
      3 ~ "democratic elections"
    ),
    .after = "electoral_category_index"
  ) |> 
  dplyr::mutate(
    election_month = stringr::str_squish(.data$election_month),
    dplyr::across(
      c(
        tidyselect::ends_with("_index"),
        tidyselect::contains("year"),
        tidyselect::ends_with("_members"),
        "parliament_chambers"
      ),
      as.integer
    ),
    dplyr::across(
      c(
        tidyselect::starts_with("is_"),
        tidyselect::starts_with("has_")
      ),
      as.logical
    )
  )

view(democracy_data)

## Checking values
unique(democracy_data$is_female_monarch)
```

# Learn one new thing
I decided to learn to make a population pyramid. These are typically used to look at number of people at different age groups and genders, living in specific places. I followed this [website](https://www.geeksforgeeks.org/how-to-create-a-population-pyramid-in-r/) as a resource to learn how to do this. 

The important part about this actually comes from the _data manipulation_, rather than code for a plot. You just make one of the groups negative, the change the axis labels so it can actually be displayed as positive. 

### Manipulate data
```{r}
democracy_data_filt <- democracy_data %>%                                 ## get data and rename
  select(year, is_female_monarch) %>%                                     ## select only columns year and gender
  filter(!is.na(is_female_monarch)) %>%                                   ## remove NAs for useful data
  mutate(monarch_gender = ifelse(is_female_monarch == "TRUE", "Female", "Male"))%>%  ## Change true/false to Female/Male
  group_by(year, monarch_gender) %>%                                      ## group by year and gender
  summarise(count = n(), .groups = 'drop') %>%                            ## get counts for each group
  mutate(count = ifelse(monarch_gender=="Female", count*(-1),count*1))    ## change female counts to negative for graph

```

### Make plot
```{r}
ggplot(democracy_data_filt, aes(x = year,                               ## call data and x, y, and fill
                                y = count, 
                                fill=monarch_gender)) +  
    geom_bar(stat = "identity") +                                       ## use bar plot
    coord_flip()+                                                       ## flip the axis
  scale_y_continuous(                                                  
    limits = c(-35, 35),                                                ## change the scales for aesthetic purposes
    breaks = seq(-40, 40, 10),                                          ## place breaks where I want them
    labels = abs(seq(-40, 40, 10))                                      ## use absolute values for female side so it's consistent
  ) + 
  scale_fill_manual(values = c("Female" = "darkblue", "Male" = "lightblue")) +  ## change the colors
   labs(title = "Comparison of Ruler Gender from 1950-2020",                    ## change names of title, axis, and legend
        x = "Year",  
        y = "Number of Rulers", 
        fill = "Gender") + 
  theme_minimal()                                                       ## minimal theme

ggsave("ruler_gender.png", width = 8, height = 6)            ## Save plot

```

