---
title: 'Tidy Tuesday Week 13/ #4'
author: "Kassandra Jensen"
date: "2024-11-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library
```{r}
library(here)
library(tidyverse)
```

# Load Data
```{r}
episode_metrics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-19/episode_metrics.csv')
```

# Visualization
```{r}
ggplot(episode_metrics, aes(x= avg_length, y= question_ratio, color = season))+
  geom_point()+
  geom_smooth(method = "lm", color = "black")+
  labs(title= "Dialogue relationships in Bob's Burgers",
       subtitle = "Relationshop between episode length and question ratio",
       x= "Average Episode Length",
       y = "Question ratio",
       color = "Season") +
  theme_minimal()
```

# One new thing I learned:
I learned how to embed a spotify link, so that it pops up with the audio of your choice. Unfortunately, in this case, the user would need to have spotify. *The markdown also must be viewed in a browser*. Now you can play a nice Bob's Burgers song in the background while enjoying your Bob's Burgers visual!


There are multiple ways to do this actually, but I found that the best was just changing the output to HTML and putting the link provided by spotify. You can also follow the directions from this (forum link)[https://forum.posit.co/t/how-do-i-embed-a-spotify-song-in-rmarkdown/116809/2] for an alternate way to do this. 

Because you can't see my code in the knitted html, I copied and pasted it. :)

"```{=html}
<iframe src="https://open.spotify.com/track/0yw7ldZLm8lUNcwDyCVh6F?si=a815bc8979a8453b" width="100%" height="380" frameBorder="0" allowtransparency="true" allow="encrypted-media" data-external="1"></iframe>
```"

```{=html}
<iframe src="https://open.spotify.com/track/0yw7ldZLm8lUNcwDyCVh6F?si=a815bc8979a8453b" width="100%" height="380" frameBorder="0" allowtransparency="true" allow="encrypted-media" data-external="1"></iframe>
```

