---
title: "Week 9 Homework"
author: "Kassandra Jensen"
date: "2024-10-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Directions: Use any dataset you would like and make two useful functions. One that is a plot and one that is not. Turn in your functions using RMarkdown with at least 2 examples per function showing its flexibility. Due, as always, at 1pm on Tuesday.

# Load Libraries
```{r}
library(tidytuesdayR)
library(ggmap)
library(tidyverse)
library(here)
library(ggspatial)
library(lme4)
```

## Use Chess Data
```{r}
tuesdata <- tidytuesdayR::tt_load('2024-10-01')

chess <- tuesdata$chess

view(chess)
colnames(chess)

```

# Function 1
## Linear Model Function
This allows you to change the response and predictor variables. 
```{r}
model_function <- function(response, predictors){                   ##add variables that will be used in function
  formula <- as.formula(paste(response, "~",                        ## use paste to take multiple variables but separate by a character 
                              paste(predictors, collapse = " + "))) ## used collapse to make it one line/ string 
  model <- lm(formula, data = chess)                                ## always use lm and data = chess 
}

## Example 1:
model_summary <- model_function("white_rating", c("turns", "time_increment"))     ## model function with response, then two predictors
model_summary

## Example 2:
model_summary2 <- model_function("white_rating", c("winner", "victory_status"))   ## model function with response and one predictor
model_summary2
```

# Function 2
## Plot function
```{r}
## unsure what to doo... could do a general one where you plug in the type of ggplot you want...
ggplot(chess, aes(x = winner, y = turns)) +
  geom_col() +
  labs(title = "Distribution of Game Turns",
       x = "Turns",
       y = "Count") +
  theme_minimal()

```

