---
title: "Week 9 Homework"
author: "Kassandra Jensen"
date: "2024-10-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Directions: Use any dataset you would like and make two useful functions. One that is a plot and one that is not. Turn in your functions using RMarkdown with at least 2 examples per function showing its flexibility. Due, as always, at 1pm on Tuesday.

# Load Libraries
```{r}
library(tidytuesdayR)
library(ggmap)
library(tidyverse)
library(here)
library(ggspatial)
library(lme4)
```

## Use Chess Data
This data was obtained by looking through the tidy tuesday repository located [here](https://github.com/rfordatascience/tidytuesday).
The specific folder containing this data can be found [here](https://github.com/rfordatascience/tidytuesday/tree/master/data/2024/2024-10-01).
```{r}
tuesdata <- tidytuesdayR::tt_load('2024-10-01')

chess <- tuesdata$chess
colnames(chess)
glimpse(chess)
view(chess)
```

# Function 1
## Linear Model Function
This allows you to change the response and predictor variables. 
Function is based off the following linear model:
    model_name <- glm(response ~ predictor 1 + predictor 2,
                      data = data_name,
                      family = "binomial")
```{r}
## First we need to convert winner into a binary so it's more of a probability outcome of each color winning because you can't have a 
## catergorical variable for a response in a glm. 

chess$winner_binary <- ifelse(chess$winner == "white", 1, 0)       ## use if else to say "IF white, then value = 1, or ELSE value = 0"

## Make function for model to predict outcome

model_function <- function(response, predictors){                   ##add variables that will be used in function
  formula <- as.formula(paste(response, "~",                        ## use paste to take multiple variables but separate by a character 
                              paste(predictors, collapse = " + "))) ## used collapse to make it one line/ string 
  model <- glm(formula, data = chess,                               ## always use glm, data = chess,
                      family = "binomial")                          ## always use family= binomial because we are using win and loss as response
}


## Example 1:
model_summary <- model_function("winner_binary", c("turns", "victory_status"))     ## model function with response, then two predictors
model_summary

## Example 2:
model_summary2 <- model_function("winner_binary", c("white_rating"))               ## model function with response and one predictor
model_summary2
```

# Function 2
## Plot function
We can use this function to showcase the results from the models from function 1.
```{r}
function2 <-function(data = chess, x, y, lines=TRUE ){     ## variables used in function
ggplot(data, aes(x = {{x}}, y = {{y}}))+                   ## formula of function, x and y can be interchanged
 geom_boxplot() +                                          ## using boxplot to showcase results
  theme_bw()                                               ## basic black and white theme
}

## I did not use the if, else statement here like we did in class because these are simple graphs

## Example 1:
function2(x = winner, y = white_rating) 

ggsave("plot1_HW9.png", width = 8, height = 6)

## Example 2:
function2(x = winner, y = turns)

ggsave("plot2_HW9.png", width = 8, height = 6)



```


