---
title: "Week 16 Lecture"
author: "Kassandra Jensen"
date: "2024-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)
```

```{r}
print(paste("The year is", 2000))
```


Let's ask R to print a statement telling us what year it is along a sequence of years.

Start with the simple code for one year
```{r}
years<-c(2015:2021)
for (i in years){ # set up the for loop where i is the index
  print(paste("The year is", i)) # loop over i
}
```
What we just did printed something over and over, but it did not save it anywhere. Let's say you want to save a new vector with all the years. To do this we need to pre-allocate space and tell R where it is going to be saved. Create an empty dataframe called year_data with columns for year and year_name .
```{r}
#Pre-allocate space for the for loop
# empty matrix that is as long as the years vector
 year_data<-tibble(year =  rep(NA, length(years)),  # column name for year
                   year_name = rep(NA, length(years))) # column name for the year name
year_data
```


Add the for loop

One line at a time. Let's first add in the column that is going to have all the names in it. Notice that I added an index i in the column name. I also am having the index go from 1:length(years), which is 1:7. I use length() because it allows me to change the number of years in the vector without having to change the for loop.
```{r}
for (i in 1:length(years)){ # set up the for loop where i is the index
  year_data$year_name[i]<-paste("The year is", years[i]) # loop over i
}
year_data
```

Fill in the year column too
```{r}
for (i in 1:length(years)){ # set up the for loop where i is the index
  year_data$year_name[i]<-paste("The year is", years[i]) # loop over year name
  year_data$year[i]<-years[i] # loop over year
}
year_data
```

#Practice
```{r}
testdata<-read_csv(here("Week15", "Lecture", "cond_data","011521_CT316_1pcal.csv"))
glimpse(testdata)

CondPath<-here("Week15", "Lecture", "cond_data")

# list all the files in that path with a specific pattern
# In this case we are looking for everything that has a .csv in the filename
# you can use regex to be more specific if you are looking for certain patterns in filenames

files <- dir(path = CondPath,pattern = ".csv")
files
```

Let's calculate the mean temperature and salinity from each file and save it
```{r}
# pre-allocate space
# make an empty dataframe that has one row for each file and 3 columns
 cond_data<-tibble(filename =  rep(NA, length(files)),  # column name for year
                   mean_temp = rep(NA, length(files)), # column name for the mean temperature
                   mean_sal = rep(NA, length(files)), # column name for the mean salinity
                   ) # column name for the year name
cond_data
```

write basic code to calculate a mean and build out
```{r}
raw_data<-read_csv(paste0(CondPath,"/",files[1])) # test by reading in the first file and see if it works
head(raw_data)
```

```{r}
mean_temp<-mean(raw_data$Temperature, na.rm = TRUE) # calculate a mean
mean_temp
```
Turn it into a for loop
```{r}
for (i in 1:length(files)){ # loop over 1:3 the number of files
}
```

Add in the loop over the raw data
```{r}
for (i in 1:length(files)){ # loop over 1:3 the number of files 
raw_data<-read_csv(paste0(CondPath,"/",files[i]))
glimpse(raw_data)
}
```
Add in the columns

First, add in the filename for each row
```{r}
for (i in 1:length(files)){ # loop over 1:3 the number of files 
raw_data<-read_csv(paste0(CondPath,"/",files[i]))
#glimpse(raw_data)
cond_data$filename[i]<-files[i]
} 
cond_data
```


```{r}
for (i in 1:length(files)){ # loop over 1:3 the number of files 
raw_data<-read_csv(paste0(CondPath,"/",files[i]))
#glimpse(raw_data)
cond_data$filename[i]<-files[i]
cond_data$mean_temp[i]<-mean(raw_data$Temperature, na.rm =TRUE)
cond_data$mean_sal[i]<-mean(raw_data$Salinity, na.rm =TRUE)
} 
cond_data
```

#simple example
```{r}
# create a vector
1:10

# for each time 1:10 make a vector of 15 random numbers based on a normal distribution
1:10 %>% # a vector from 1 to 10 (we are going to do this 10 times) %>% # the vector to iterate over
  map(rnorm, n = 15) # calculate 15 random numbers based on a normal distribution in a list

#Calculate the mean from each list

1:10 %>% # a vector from 1 to 10 (we are going to do this 10 times) %>% # the vector to iterate over
  map(rnorm, n = 15)  %>% # calculate 15 random numbers based on a normal distribution in a list 
  map_dbl(mean) # calculate the mean. It is now a vector which is type "double"

```

# Make your own function
```{r}
1:10 %>% # list 1:10
  map(function(x) rnorm(15, x)) %>% # make your own function
  map_dbl(mean)

#Use a formula when you want to change the arguments within the function

1:10 %>%
  map(~ rnorm(15, .x)) %>% # changes the arguments inside the function
  map_dbl(mean)
```

# More example
```{r}
CondPath<-here("Week16","cond_data")
files <- dir(path = CondPath,pattern = ".csv")
files
```
Or, we can get the full file names in one less step by doing this.
```{r}
files <- dir(path = CondPath,pattern = ".csv", full.names = TRUE)
#save the entire path name
files
```

Next, read in the files using map instead of a for loop while retaining the filename as a column.
```{r}
data<-files %>%
  set_names()%>% # set's the id of each list to the file name
  map_df(read_csv,.id = "filename") # map everything to a dataframe and put the id in a column called filename
data
```
Calculate means
Now we have a regular dataframe and we can calculate means in the way we already know how! group_by filename and use summarize
```{r}
data<-files %>%
  set_names()%>% # set's the id of each list to the file name
  map_df(read_csv,.id = "filename") %>% # map everything to a dataframe and put the id in a column called filename
  group_by(filename) %>%
  summarise(mean_temp = mean(Temperature, na.rm = TRUE),
            mean_sal = mean(Salinity,na.rm = TRUE))
data
```


