---
title: "Good plot/ bad plot"
author: "Kassandra Jensen"
date: "2024-10-24"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      messge = FALSE)
```

# Directions:
Due October 29

Create two version of similar plots, using any dataset you choose. You must create the plots using ggplot2, post-processing with Adobe Illustrator or similar is not allowed.
Please submit your homework by posting a markdown document containing the plots, the code you used to create them, and your discussion of the plot within the markdown doc to github. Data also needs to be submitted.

# Load libraries
```{r}
library(tidyverse)
library(here)
library(glue)
```

# Load data:

The data for this assignment came from the tidy tuesday posted on 12/19/2022. A link to the repository can be found [here](https://github.com/rfordatascience/tidytuesday/tree/master/data/2022/2022-12-20).

There are three separate data sheets that were downloaded from the repository, however only two were used in this assignment. 

This data describes weather forecasts and conditions across the United States over a period of 16 months, from January 2021-June2022.
```{r}
weather_forecasts <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-12-20/weather_forecasts.csv')
outlook_meanings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-12-20/outlook_meanings.csv')

view(weather_forecasts)
view(outlook_meanings)

## Saved data and uploaded to repository - do not need to keep saving 
#write.csv(weather_forecasts, file = "weather_forecasts.csv", row.names = FALSE)
#write.csv(outlook_meanings, file = "outlook_meanings.csv", row.names = FALSE)
```

# Clean Data
```{r}
weather_forecasts_HI <- weather_forecasts%>%          ## Call data
  filter (state == "HI") %>%                          ## Only Hawai‘i
  drop_na(forecast_outlook)                           ## Get rid or NAs in column

## Check for which types of forecast outlooks there are 
unique(weather_forecasts_HI$forecast_outlook)

## Rename everything to unabbreviated name 
weather_forecasts_HI <- weather_forecasts_HI %>%                                                    
  mutate(forecast_outlook = if_else(forecast_outlook == "SUNNY", "Sunny", forecast_outlook)) %>%
  mutate(forecast_outlook = if_else(forecast_outlook == "MOCLDY", "Mostly Cloudy", forecast_outlook))%>%
  mutate(forecast_outlook = if_else(forecast_outlook == "PTCLDY", "Partly Cloudy", forecast_outlook))%>%
  mutate(forecast_outlook = if_else(forecast_outlook == "SHWRS", "Rain Showers", forecast_outlook))%>%
  mutate(forecast_outlook = if_else(forecast_outlook == "CLOUDY", "Cloudy", forecast_outlook))%>%
  mutate(forecast_outlook = if_else(forecast_outlook == "WINDY", "Windy", forecast_outlook))%>%
  mutate(forecast_outlook = if_else(forecast_outlook == "TSTRMS", "Thunderstorms", forecast_outlook))
```

# Bad plot
## Data manipulation
```{r}
weather_forecasts_HI_counts<- weather_forecasts_HI %>% 
  group_by(forecast_outlook) %>%
  summarise(count = n())

## View to make sure it's what I want
view(weather_forecasts_HI_counts)
```

##### This is a bad plot because:
- Missing key variables that make this data useful, like time or location

- Color are distracting

- Title does not explain subject of plot approriately 

- Font size too big in plot and is being cut off

- Inappropriate use of emojis
```{r}
ggplot(weather_forecasts_HI_counts, aes(x = reorder(forecast_outlook, -count), y = 1, fill = count)) +
  geom_tile(color = "magenta", size = 2) + 
  geom_text(aes(label = count), size = 23, color = "yellow", angle = 90) +  
  scale_fill_gradient(low = "limegreen", high = "red") +  
  labs(title = expression(paste("😎 How much weather?? 🌧")),
                                x = "Forecast Outlook", y = "") +
  theme_void() + 
  theme(
    plot.title = element_text(size = 15, face = "bold.italic", hjust = 0.5, color = "blue"),  
    legend.position = "top", 
    panel.grid.major = element_line(size = 1, color = "cyan"),  
    panel.grid.minor = element_line(size = 1, color = "purple"),  
    axis.text.x = element_text(size = 14, face = "bold", color = "orange", angle = 45)  
  ) +
  coord_fixed(ratio = 1.5)  

ggsave(here("Good_plot_bad_plot","Output","badplot.png"))
```
# Good plot
### Data manipulation
```{r}
weather_forecasts_HI <- weather_forecasts_HI %>%   ## call data
  mutate(month_year = floor_date(date, "month"))   ## round data to nearest month 

forecast_counts <- weather_forecasts_HI %>%        ## call data
  group_by(month_year, forecast_outlook) %>%       ## group by date and forecast because that's all we care about
  summarise(count = n()) %>%                       ## n() counts the unique observations of what I put in groupby() 
  ungroup()  

```

### Plot
##### This is a good plot because:
- Clear title and subtitle showcasing location and what the plot is trying to represent

- Helpful x and y-axis

- Appropriate font size 

- Colors are not distracting and help convey denities

- Heat map conveys meaningful densities overtime for comparisons

- Data is sourced
```{r}
ggplot(forecast_counts, aes(x = month_year,                             ## call data, put month_year as x axis
                            y = forecast_outlook)) +                    ## forecast as y- axis
  geom_tile(aes(fill = count), color = "white") +                       ## use geom_tile for heat map
  scale_fill_gradient(low = "lightblue", high = "darkblue") +           ## pretty colors
  labs(title = "Temporal Analysis of Weather Forecasts",                ## titles
       subtitle = "1.5 Years of Honolulu Weather Data",
       caption = "Data source: https://github.com/rfordatascience/tidytuesday/tree/master/data/2022/2022-12-20",
       x = "Month-Year",
       y = "Forecast",
       fill = "Count") +
  theme_minimal() +                                                     ## minimal theme
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(here("Good_plot_bad_plot","Output","goodplot.png"))
```

